Comparative Evaluation of Spark and Flink Stream Processing

### 2017-03-03 paper/paper.tex
[paper] updated sector change detection in Spark

\par The Spark Streaming implementation of trajectory statistics computation reads the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream} that represents the stream of position messages, in order to scale up and have parallel Kafka stream receivers, a multiple instances of \texttt{DStream} must be created and combined together using the \texttt{union} operation. The output stream of the \texttt{union} is further processed using \texttt{mapToPair}, to map each position message to tuple of aircraft ID and position message. Then, the irrelevant tuples are filtered based on the position message type using  a \texttt{filter} transformation. Afterward, the tuples that share the same aircraft ID are groped together to construct the stream of trajectories (tuples of Id and list of positions), by applying the \texttt{groupByKey} operation. Since the processing  model of Spark Streaming is micro-batch based and stateless, \texttt{updateStateByKey} operation must be used to maintain the state of the trajectories stream. In the context of this workload, the state is the aggregated statistics of trajectory, so the statistics computation is preformed within the custom \texttt{updateStateByKey} function that continuously applied for every batch.
 The state update function uses the last position of trajectory from previous batch to aggregate and compute statistics quantities to each trajectory's position in the new data batch, and the new positions list is kept as the new state for the trajectory. Inside the state update function, the new positions must be sorted to preserve the correct order, since the tuples after \texttt{groupByKey} operation are shuffled across the cluster's  nodes and sent to the state update function.

@@

The detail of Flink's implementation of statistics computation workload as follows. First, a \texttt{FlinkKafkaConsumer} is used to read the Kafka stream that constructs \texttt{DataStream} instance, moreover, it implicitly runs multiple parallel Kafka stream consumers, which makes the Kafka partitions balanced between the instances. Second, the stream records (i.e., ADS-B messages) are parsed and transformed to tuples of ID and position message using a configured \texttt{map} transformation. Third, a \texttt{filter} function is used to filter irrelevant tuples based on the message type. Fourth, the result tuples after the filtration steps are grouped with similar ID using \texttt{keyBy} operator, to constructs the stream of trajectories. Finally, 
a \texttt{reduce} transformation on the trajectories tuples is used to calculate the statistics for new received trajectory's position, utilizing the aggregated statistics of previous position, while old position is discarded and the new record is used to aggregate the statistics.

@@

The goal of sector change detection workload is to detect the  enter or leaving of aircraft from one air sector to other, by processing the real-time stream of trajectories. Given that the dataset of sectors (i.e., polygons) is available as reference to assign the corresponding sector for certain aircraft's position. In this workload, we test the aspect of providing the  global model of data (sectors dataset) in streaming processing workload. Moreover, the preserving od stream's records order and consuming of parallel Kafka streams points were covered.

@@

 We implemented the sector change detection workload in Spark by constructing the trajectories stream and assigning the sectors for each position in all trajectories, then a change between the assigned sectors of two consecutive aircraft's positions is marked as alert case. First, a multiple instances of \texttt{DStream} are created using \texttt{KafkaUtils.createDirectStream} to read in  parallel from Kafka, after  all input data streams are combined using \texttt{union} function. Second, the records (i.e., ADS-B messages) of the resulted union stream are parsed and mapped to tuples of (aircraft ID, position message) using defined \texttt{mapToPair} transformation. Third, the unrelated  tuples are filtered based on the message type through \texttt{filter} operation. Fourth, the trajectories stream is constructed by combining all tuples with the same ID using the \texttt{groupByKey} function. Then, to manage and aggregate the state of trajectories stream between the mini-batches a configured \texttt{updateStateByKey} function is applied,  the corresponding sector are assigned, the old positions of certain trajectory from previous batch are discarded, but the last position of old batch is kept to be used alongside with the new positions.
 The sector data set is provided using the \texttt{Broadcast} shared value feature in Spark that distributes the  data set to all cluster nodes efficiently. Finally, the detection of sector change is done by extra filtering transformation using \texttt{filter}, by checking the sector of the positions list in each tuple.

>>>>

\par The implementation of trajectory statistics computation in Spark Streaming, first, it reads the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream} that represents the stream of position messages, in order to scale up and have parallel Kafka stream receivers, a multiple instances of \texttt{DStream} must be created and combined together using the \texttt{union} operation. The output stream of the \texttt{union} is further processed using \texttt{mapToPair}, to map each position message to tuple of aircraft ID and position message. Then, the irrelevant tuples are filtered based on the position message type using  a \texttt{filter} transformation. Afterward, the tuples that share the same aircraft ID are groped together to construct the stream of trajectories (tuples of Id and list of positions), by applying the \texttt{groupByKey} operation. Since the processing  model of Spark Streaming is micro-batch based and stateless, \texttt{updateStateByKey} operation must be used to maintain the state of the trajectories stream. In the context of this workload, the state is the aggregated statistics of trajectory, so the statistics computation is preformed within the custom \texttt{updateStateByKey} function that continuously applied for every batch.
 The state update function uses the last position of trajectory from previous batch to aggregate and compute statistics quantities to each trajectory's position in the new data batch, and the new positions list is kept as the new state for the trajectory. Inside the state update function, the new positions must be sorted to preserve the correct order, since the tuples after \texttt{groupByKey} operation are shuffled across the cluster's  nodes and sent to the state update function.

@@

The details of Flink's implementation of statistics computation workload as follows. First, a \texttt{FlinkKafkaConsumer} is used to  construct a \texttt{DataStream} instance that represent the Kafka stream of aircrafts positions, it implicitly runs multiple parallel Kafka stream consumers, which makes the Kafka partitions balanced between the instances. Second, the stream records (i.e., ADS-B messages) are parsed and transformed to tuples of ID and position message using a configured \texttt{map} transformation. Third, a \texttt{filter} function is used to filter irrelevant tuples based on the message type. Fourth, the resulted tuples after the filtration transformation are grouped based on the tuple's ID using \texttt{keyBy} operator that constructs the \texttt{KeyedStream} of trajectories. Finally, 
a \texttt{reduce} transformation on the  \texttt{KeyedStream} is used to calculate the statistics for each new coming trajectory's position, by utilizing the aggregated statistics of previous position, while old position is discarded and the new position with attached computed statistics is used to aggregate the statistics of a certain trajectory.

@@

The goal of sector change detection workload is to detect the  enter or leaving of aircraft from one air sector to other, by processing the real-time stream of aircraft positions. Given that the dataset of sectors (i.e., polygons) is available as reference to assign the corresponding sector for certain aircraft's position. In this workload, we test the aspect of providing the  global model of data (sectors dataset) in streaming processing workload. Moreover, the preserving of the stream's records order and consuming of parallel Kafka streams points are also covered.

@@

 We implemented the sector change detection workload in Spark Streaming by constructing the trajectories stream and assigning the sectors for each position of a trajectory, then a change between the assigned sectors of two consecutive aircraft's positions is marked  as  an transition of sector for certain trajectory. First, a multiple instances of \texttt{DStream} are created using \texttt{KafkaUtils.createDirectStream} to read in  parallel from Kafka, after  all input data streams are combined using \texttt{union} function. Second, the records (i.e., ADS-B messages) of the resulted union stream are parsed and mapped to tuples of (aircraft ID, position message) using defined \texttt{mapToPair} transformation. Third, the unrelated  tuples are filtered based on the message type through a \texttt{filter} operation. Fourth, the trajectories stream is constructed by combining all tuples with the same ID using the \texttt{groupByKey} function. Then, to manage and aggregate the state of trajectories stream between the mini-batches a configured \texttt{updateStateByKey} function is applied. The state update function assign the corresponding sector for each new position, the old positions of a certain trajectory from previous batch are discarded, but only the last position of old batch is kept to be used alongside with the new positions. In addition, the new positions are sorted to maintain the correct order.
 The sector data set is provided using the built-in \texttt{Broadcast} feature that distributes a shared value for  all cluster nodes in efficient matter. Finally, the detection of sector change is done by an extra filtering transformation using a \texttt{filter} operator, by checking the sectors of consecutive positions in a trajectory to find the ones with a change in the sector assignment.
